---
title: "lab-07-simpsons.Rmd"
author: "Ibtesam Faleh 2201002692 "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer:The data is observational as the description states that is based on age, smoking, and mortality, which are all observable events and not produced via experiments.

2.  
```{r observations}
nrow(Whickham)
```

Your answer;Represent recorded participants"age",smoking status at baseline.

3.  

```{r number_variables}
names(Whickham)
```

Your answer:  age (Numerical),smoker and outcome are categorical.

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)       
unique(Whickham$age)
```

Your answer: Using the 'unique()' function on the 3 variables we could see that"outcome" only takes Alive or Dead value,which makes it categorical non ordinal.."smoker"only takes yes or no,which also makes it categorical non ordinal.Age is numerical contincus data.
one of the best ways to visualise  categorical data is through the use of bar charts.

```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +

 geom_bar()
```


```{r plot_smoker}
ggplot(Whickham, aes(x = smoker)) +

 geom_bar()
```

```{r plot_age}
ggplot(Whickham, aes(x = age)) +
  geom_boxplot()
```


4. I expect the health will be worser and may be the person will be died after while, if he keeping smoke.

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

Knit, commit, and push to github.

5. 

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
```

6.

```{r age_cat}

```

7. 

```{r plot_smokeVSincome_3}

```

Knit, commit, and push to github.